---
title: "Final Project - JHU Practical Machine Learning"
author: "George Ren"
date: "10/1/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem Statement

In this exercise, we are trying to predict how well people are lifting barbells using measurement data from accelerometers on the belt, forearm, arm, and dumbell.  Refer to <http://groupware.les.inf.puc-rio.br/har> for more details.


### Step 0: Read in the data
We begin with downloading the data files from the Internet and save them as csv files locally.
```{r, eval=FALSE}
rawTrain <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", header=TRUE)
rawTest <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", header=TRUE)
write.csv(rawTrain, "rawTrain.csv")
write.csv(rawTest, "rawTest.csv")
```

### Step 1: Data Cleaning and Preprocessing
Once the data is downloaded successfully, we then take a moment to clean and preprocess the data.

First, we identify variables with mostly zero values and remove them from both the training and testing datasets.
```{r, eval=FALSE}
# Remove first row which is essentially row numbers
rawTrain <- rawTrain[, -1]
rawTest <- rawTest[, -1]

nzv <- nearZeroVar(rawTrain, saveMetrics = TRUE)
training <- rawTrain[, nzv$zeroVar==FALSE & nzv$nzv==FALSE]
testing <- rawTest[, nzv$zeroVar==FALSE & nzv$nzv==FALSE]
```

Second, we find variables with high missing rate and remove them from both datasets.
```{r}
missPerc <- colMeans(is.na(training))       # calculate NA percentage
training <- training[, missPerc <= 0.8]

missPerc <- colMeans(is.na(testing))       # calculate NA percentage
training <- testing[, missPerc <= 0.8]
```




## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
